NumPy---Array Programming in Python

### 2020-01-26 body.tex
Improve ecosystem figure

The brevity and expressive power of NumPy's array programming syntax combined with 
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa} makes interactive data exploration fluid and simple.
It is common for NumPy users to work interactively with their data,
discovering through experimentation bits of code that they stitch together into
imperative or functional programs.
The scientific Python ecosystem (Fig.~\ref{fig:ecosystem}) builds on top of
this solid foundation with core (e.g., SciPy and Matplotlib) libraries,
widely used technique-specific
libraries\cite{pedregosa2011scikit,vanderwalt2014scikit,SciPyProceedings_11},
and several domain-specific projects and libraries
\cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,2018EGUGA..2012146H}.
Exposing these simple array programming primitives, as well as the surrounding ecosystem of
tools, in Python---an interpreted, general purpose language---creates an ideal environment for
interactive, exploratory data analysis where users may iteratively inspect,
manipulate, and visualize their data.
This rich environment has made the scientific Python ecosystem popular for scientific
research.

@@

For example, the \code{eht-imaging} library developed by the Event Horizon Telescope 
collaboration relies on many components of the scientific Python ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image 
reconstruction.
SciPy is used to provide tools used in general image processing such as 
filtering and image alignment, while scikit-image provides higher-level
functionality such as edge filters and Hough transforms.
SciPy provides fundamental algorithms for scientific computing on arrays
including many mathematical, scientific, and engineering routines.
SciPy and NumPy are closely related and were developed together often
by the same people.
scikit-image is a popular technique-specific library,
which vastly extends the basic image processing tools provided by SciPy.
Other components of SciPy are used throughout the library, such as the
\code{scipy.optimize} package for handling general optimization tasks.
NetworkX \cite{SciPyProceedings_11} (the standard package for complex
network analysis) is used for analysis of consistency among image
comparisons.
Astropy \cite{astropy:2013, astropy:2018} (a community-developed astronomy
library) is used throughout the library,
primarily for I/O from standard astronomical file formats and time/coordinate
transformations.
Matplotlib (the core library for generating publication-ready figures) is
used for visualizing data throughout the analysis pipeline,
including the generation of the final image of the black hole.
% N.B. - most of this comes from introspecting the Image class defined in
% ehtim/image.py and grepping the dependencies in requirements.txt

Tools for analyzing data from gravitational wave observatories such as LIGO
and Virgo are provided by the \code{pycbc} package, which was used in the first
detection of gravitational waves \cite{abbott2016observation} and in 
on-going analysis of data from LIGO and Virgo.
\code{pycbc} make extensive use of the scientific Python ecosystem.
The time-series data from the interferometers are stored in NumPy arrays,
while the \code{scipy.signal} is used to construct and apply filters to the
data.
Matplotlib is used to visualize data at all points along the analysis chain
including raw data from the instrumentation and the time-frequency 
visualization of the ``chirps'' that result from mergers of compact binaries
such as black holes\cite{abbott2016observation} and
neutron stars\cite{abbott2017multi}.

>>>>

The brevity and expressive power of NumPy's array programming syntax combined with 
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa} makes interactive data exploration fluid and simple.
It is common for NumPy users to work interactively with their data,
discovering through experimentation bits of code that they stitch together into
imperative or functional programs.
SciPy and Matplotlib are tightly integrated with NumPy to expand the suite of
standard algorithms and provide basic plotting capabilities.
SciPy provides fundamental algorithms for scientific computing on arrays
including many mathematical, scientific, and engineering routines.
Matplotlib generates publication-ready figures and visualizations.
The scientific Python ecosystem (Fig.~\ref{fig:ecosystem}) builds on top of
this solid foundation providing many, widely used technique-specific
libraries\cite{pedregosa2011scikit,vanderwalt2014scikit,SciPyProceedings_11}
and several domain-specific projects and libraries
\cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,2018EGUGA..2012146H}.
Exposing these simple array programming primitives, as well as the surrounding ecosystem of
tools, in Python---an interpreted, general purpose language---creates an ideal environment for
interactive, exploratory data analysis where users may iteratively inspect,
manipulate, and visualize their data.
This rich environment has made the scientific Python ecosystem popular for scientific
research.

@@

For example, the \code{eht-imaging} library developed by the Event Horizon Telescope 
collaboration relies on many components of the scientific Python ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image 
reconstruction.
SciPy is used to provide tools used in general image processing such as 
filtering and image alignment, while scikit-image provides higher-level
functionality such as edge filters and Hough transforms.
scikit-image is a popular technique-specific library,
which vastly extends the basic image processing tools provided by SciPy.
Other components of SciPy are used throughout the library, such as the
\code{scipy.optimize} subpackage for handling general optimization tasks.
NetworkX \cite{SciPyProceedings_11} (the standard package for complex
network analysis) is used for analysis of consistency among image
comparisons.
Astropy \cite{astropy:2013, astropy:2018} (a community-developed astronomy
library) is used throughout the library,
primarily to handle standard astronomical file formats and time/coordinate
transformations.
Matplotlib is
used for visualizing data throughout the analysis pipeline,
including the generation of the final image of the black hole.
% N.B. - most of this comes from introspecting the Image class defined in
% ehtim/image.py and grepping the dependencies in requirements.txt

Tools for analyzing data from gravitational wave observatories such as LIGO
and Virgo are provided by the \code{pycbc} package, which was used in the first
detection of gravitational waves \cite{abbott2016observation} and in 
on-going analysis of data from LIGO and Virgo.
\code{pycbc} make extensive use of the scientific Python ecosystem.
The time-series data from the interferometers are stored in NumPy arrays,
while the \code{scipy.signal} subpackage is used to construct and apply filters to the
data.
Matplotlib is used to visualize data at all points along the analysis chain
including raw data from the instrumentation and the time-frequency 
visualization of the ``chirps'' that result from mergers of compact binaries
such as black holes\cite{abbott2016observation} and
neutron stars\cite{abbott2017multi}.
